\documentclass[aspectratio=169, table]{beamer}

\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{polar}
\usetikzlibrary{arrows.meta, positioning, calc}

\usetheme{Pradita}


\usepackage{listings}
\lstdefinestyle{SqlStyle}{
	language=SQL,
	basicstyle=\ttfamily\footnotesize,
	morekeywords={REAL, TEXT, REFERENCES},
	keywordstyle=\color{blue},
	commentstyle=\color{gray},
	stringstyle=\color{red},
	breaklines=true,
	showstringspaces=false,
	tabsize=2,
	captionpos=b,
	numbers=left,
	numberstyle=\tiny\color{gray},
	frame=lines,
	backgroundcolor=\color{lightgray!10},
	comment=[l]{//},
	morecomment=[s]{/*}{*/},
	commentstyle=\color{gray}\ttfamily,
	string=[s]{'}{'},
	morestring=[s]{"}{"},
	%	stringstyle=\color{teal}\ttfamily,
	%	showstringspaces=false
}

\lstdefinelanguage{bash} {
	keywords={},
	basicstyle=\ttfamily\small,
	keywordstyle=\color{blue}\bfseries,
	ndkeywords={iex},
	ndkeywordstyle=\color{purple}\bfseries,
	sensitive=true,
	commentstyle=\color{gray},
	stringstyle=\color{red},
	numbers=left,
	numberstyle=\tiny\color{gray},
	breaklines=true,
	frame=lines,
	backgroundcolor=\color{lightgray!10},
	tabsize=2,
	comment=[l]{\#},
	morecomment=[s]{/*}{*/},
	commentstyle=\color{gray}\ttfamily,
	stringstyle=\color{purple}\ttfamily,
	showstringspaces=false
}

% Define Java language style for listings
\lstdefinestyle{JavaScript}{
	language=Java,
	basicstyle=\ttfamily\footnotesize,
	keywordstyle=\color{blue},
	commentstyle=\color{gray},
	stringstyle=\color{red},
	breaklines=true,
	showstringspaces=false,
	tabsize=4,
	captionpos=b,
	numbers=left,
	numberstyle=\tiny\color{gray},
	frame=lines,
	backgroundcolor=\color{lightgray!10},
	comment=[l]{//},
	morecomment=[s]{/*}{*/},
	commentstyle=\color{gray}\ttfamily,
	string=[s]{'}{'},
	morestring=[s]{"}{"},
	%	stringstyle=\color{teal}\ttfamily,
	%	showstringspaces=false
}

\title{\Huge Data Monetization \& \\
	\vspace{10pt}
	Value Realization}
\subtitle{IT140704 - Big Data for Business}
%\date[Serial]{Penggunaan Large Language Model untuk Pengajaran}
\author{\textbf{Alfa Yohannis}}
\begin{document}
	
	\frame{\titlepage}
	
	
	\begin{frame}[fragile]
		\frametitle{Contents}
		\vspace{20pt}
		\begin{columns}[t]
			\column{0.5\textwidth}
			\tableofcontents[sections={1-5}]
			
			\column{0.5\textwidth}
			\tableofcontents[sections={6-20}]
		\end{columns}
	\end{frame}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Do you know how\\to monetize Big Data?}}
	\end{frame}
	
	\section{Introduction}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Introduction}}
	\end{frame}
	
	\begin{frame}{Introduction}
		\vspace{20pt}
		\begin{itemize}
			\item Exponential growth of data (volume, velocity, variety) makes it a critical strategic asset.  
			\item Data is non-rival, replicable, and increases in value when combined with other sources.  
			\item Monetization involves direct revenues (products, services) and indirect value (efficiency, decision-making, customer experience).  
			\item Organizations treat data as intangible capital, comparable to intellectual property or brand equity.  
			\item Policymakers stress economic potential but also highlight governance, privacy, and fair distribution challenges.  
			\item This chapter introduces data monetization within the framework of the Big Data Value Chain.  
		\end{itemize}
	\end{frame}
	
	\section{Data as an Economic Asset}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Data as an Economic Asset}}
	\end{frame}
	
	\begin{frame}{Data as an Economic Asset (1/2)}
		\vspace{20pt}
		\begin{itemize}
			\item Data is a non-rival, intangible asset: using it does not diminish value, and reuse often increases utility.  
			\item Key characteristics:  
			\begin{enumerate}
				\item Scalability and low marginal replication cost.  
				\item Context-dependent value—raw data needs processing and analysis.  
				\item Dynamic value shaped by technology, regulation, and markets.  
			\end{enumerate}
			\item Valuation approaches: cost-based, market-based, and revenue-based methods.  
			\item Firms like Microsoft and Google increasingly treat data as balance sheet assets.  
		\end{itemize}
	\end{frame}
	
	\begin{frame}{Data as an Economic Asset (2/2)}
		\vspace{20pt}
		\begin{itemize}
			\item Raw data has limited intrinsic value—true value emerges through transformation into insights.  
			\item The DIKW model highlights progression: Data → Information → Knowledge → Wisdom.  
			\item Advanced analytics, machine learning, and AI extract patterns and predictions from raw data.  
			\item Culture, governance, and leadership are critical for realizing business impact.  
			\item Data-driven firms report superior decision-making, innovation speed, and customer experience.  
			\item Unlocking value requires holistic alignment of technology, governance, and organizational readiness.  
		\end{itemize}
	\end{frame}
	
	\section{Economic and Business Value of Data}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Economic and Business Value of Data}}
	\end{frame}
	
	\begin{frame}{Direct Value (Revenue Creation)}
		\vspace{20pt}
		\begin{itemize}
			\item Data can be monetized directly via sales, licensing, subscriptions, or embedding into new offerings.  
			\item Key example: targeted advertising models of Google and Facebook.  
			\item Financial services use predictive analytics to tailor credit products.  
			\item Mobility firms (Uber, Grab) leverage geolocation for pricing and platform revenue.  
			\item Data marketplaces and exchanges enable cross-industry trading.  
			\item Success depends on data quality, regulatory compliance, and user trust.  
		\end{itemize}
	\end{frame}
	
	\begin{frame}{Indirect Value (Efficiency and Cost Reduction)}
		\vspace{20pt}
		\begin{itemize}
			\item Data enhances operations by embedding analytics into processes.  
			\item Predictive maintenance reduces downtime in manufacturing and logistics.  
			\item Supply chain analytics optimize procurement, inventory, and distribution.  
			\item Healthcare: electronic health records improve diagnostics and care planning.  
			\item Finance: real-time fraud detection prevents major losses.  
			\item Indirect value is often more sustainable, fueling continuous operational excellence.  
		\end{itemize}
	\end{frame}
	
	\begin{frame}{Strategic Value (Competitive Advantage)}
		\vspace{20pt}
		\begin{itemize}
			\item Beyond revenue and efficiency, data drives long-term competitive advantage.
			\item Strategic value lies in unique capabilities: turning data into insights, innovation, and customer-oriented offerings.
			\item Netflix/Amazon: behavioral data enables personalization, loyalty, and reduced churn.
			\item IoT in supply chains: enhances responsiveness and reliability.
			\item Data-driven firms innovate faster, spotting trends and adapting quickly.
			\item Platforms like Google/Alibaba build network effects, creating barriers to entry.
			\item Treating data as a strategic capability allows firms to shape industries and capture disproportionate value.
		\end{itemize}
	\end{frame}
	
	\section{Business Models for Data Monetization}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Business Models for Data Monetization}}
	\end{frame}
	
	\begin{frame}{Subscription and Licensing Models}
		\begin{itemize}
			\item Subscription: recurring fees (monthly/annual) for continuous access to datasets, analytics platforms, or APIs.  
			\item Licensing: limited rights to use specific datasets or tools for a defined scope or period.  
			\item Common in finance (Bloomberg, Reuters) where accuracy and credibility drive value.  
			\item Hybrid SaaS models combine software tools with predictive dashboards or industry benchmarks.  
			\item Strengths: stable revenue and long-term customer ties.  
			\item Challenges: pricing strategy, IP protection, and preventing unauthorized redistribution.  
		\end{itemize}
	\end{frame}
	
	\begin{frame}{Advertising and Targeted Marketing}
		\begin{itemize}
			\item Leverages behavioral, demographic, and preference data for personalized ads.  
			\item Platforms like Google, Facebook, TikTok monetize massive user interaction datasets.  
			\item E-commerce players (Amazon, Shopee) use transaction history for contextual advertising.  
			\item Key strength: scalability – larger user base means higher ad ecosystem value.  
			\item Challenges: regulatory limits (GDPR, CCPA), consumer trust, and privacy ethics.  
			\item Sustainability depends on balancing monetization with transparency and user rights.  
		\end{itemize}
	\end{frame}
	
	\begin{frame}{Data-as-a-Service (DaaS)}
		\begin{itemize}
			\item Evolution of SaaS: on-demand access to curated data via cloud APIs or marketplaces.  
			\item Reduces technical barriers – firms focus on insights, not infrastructure.  
			\item Providers: AWS Data Exchange, Azure Data Share, Google Cloud Marketplace.  
			\item Expands data sharing across industries, democratizing access and innovation.  
			\item Risks: data quality, provenance, privacy compliance, and misuse.  
			\item Governance and trust are critical to long-term sustainability of DaaS ecosystems.  
		\end{itemize}
	\end{frame}

\begin{frame}{Marketplace and Platform Models}
	\vspace{20pt}
	\begin{itemize}
		\item Marketplace: facilitates data exchange across parties via APIs and governance.  
		\item Examples: AWS Data Exchange, Snowflake Marketplace, Dawex (health, finance, mobility datasets).  
		\item Platform: multi-sided ecosystems adding analytics, integration, and reputation systems.  
		\item Value comes from \textbf{network effects}: more participants increase overall utility.  
		\item Challenges: interoperability standards, strict privacy/ethics, and reducing information asymmetry.  
		\item With regulation and trust, marketplaces can position data as a \textbf{global tradable commodity}, comparable to energy or finance.  
	\end{itemize}
\end{frame}

\begin{frame}{Hybrid and Emerging Models}
	\vspace{20pt}
	\begin{itemize}
		\item Hybrid models combine multiple approaches (e.g., SaaS subscription + external data marketplace).  
		\item Telecoms: license mobility data to third parties while using the same data for targeted ads.  
		\item Emerging trends:  
		\begin{enumerate}
			\item Data Cooperatives – collective management and fair benefit-sharing.  
			\item Blockchain/Data Tokenization – secure ownership and transparent trading.  
			\item AI-as-a-Service with Data Bundling – packaged data + trained AI models.  
		\end{enumerate}
		\item Advantages: flexibility, diversified revenue.  
		\item Challenges: governance, cross-border regulation, transparency, and trust.  
	\end{itemize}
\end{frame}

\section{Big Data Value Chain}

\begin{frame}{\hfill}
	\centering
	\Huge{\textbf{Big Data Value Chain}}
\end{frame}

\begin{frame}{Stages of the Big Data Value Chain}
	\vspace{15pt}
	\begin{itemize}
		\item Concept: raw data moves through stages to create value.  
		\item Typical stages:  
		\begin{enumerate}
			\item Data Acquisition $\rightarrow$ collect from structured/unstructured sources.  
			\item Data Storage/Integration $\rightarrow$ cloud, lakes, warehouses.  
			\item Data Processing/Cleansing $\rightarrow$ ETL, error removal, normalization.  
			\item Data Analytics/Interpretation $\rightarrow$ statistics, ML, AI methods.  
			\item Visualization/Presentation $\rightarrow$ dashboards, reports, storytelling.  
			\item Value Creation/Utilization $\rightarrow$ revenue, efficiency, social impact.  
		\end{enumerate}
		\item Value emerges through structured transformation, not raw data.  
	\end{itemize}
\end{frame}

\begin{frame}{Linking Value Chain to Monetization}
	\vspace{15pt}
	\begin{itemize}
		\item Each stage links to monetization models:  
		\begin{enumerate}
			\item Acquisition $\rightarrow$ Data-as-a-Service (APIs, marketplaces).  
			\item Storage/Integration $\rightarrow$ subscription or licensing.  
			\item Processing $\rightarrow$ premium “clean data” or ETL tools.  
			\item Analytics $\rightarrow$ subscription analytics platforms.  
			\item Visualization $\rightarrow$ premium dashboards, contextual ads.  
			\item Utilization $\rightarrow$ products/services, recommender systems, ads.  
		\end{enumerate}
		\item Full-chain management $\rightarrow$ diverse, sustainable revenue streams.  
		\item Value chain $\rightarrow$ key enabler for data monetization.  
	\end{itemize}
\end{frame}

\section{Value Realization Strategies}

\begin{frame}{\hfill}
	\centering
	\Huge{\textbf{Value Realization Strategies}}
\end{frame}

\begin{frame}{Internal Optimization and Process Efficiency}
	\vspace{15pt}
	\begin{itemize}
		\item Data used not for sale, but to improve internal operations.  
		\item \textbf{Operational efficiency} $\rightarrow$ IoT sensors enable predictive maintenance, reducing downtime and costs.  
		\item \textbf{Process optimization} $\rightarrow$ machine learning supports anomaly detection and “what-if” analysis for better decisions.  
		\item \textbf{Employee productivity} $\rightarrow$ ERP/CRM data + dashboards allow real-time KPI monitoring and corrective action.  
		\item \textbf{Quality control} $\rightarrow$ historical data predicts defects, enabling early detection and prescriptive solutions.  
		\item Value comes from internalization of benefits, boosting competitiveness.  
	\end{itemize}
\end{frame}

\begin{frame}{\LARGE{New Products, Services, and Customer Experience}}
	\vspace{15pt}
	\begin{itemize}
		\item Data drives \textbf{innovation and customer-centric strategies}.  
		\item \textbf{Data-driven products} $\rightarrow$ connected cars with predictive navigation, real-time monitoring, subscription mobility.  
		\item \textbf{Smart services \& personalization} $\rightarrow$ e-commerce and banking use recommendations and predictive offers.  
		\item \textbf{Customer experience} $\rightarrow$ social media analytics track sentiment, improve loyalty and satisfaction.  
		\item \textbf{Ecosystem services} $\rightarrow$ health apps linking patients, doctors, insurers.  
		\item Outcome: stronger loyalty, unique value propositions, new revenue streams.  
	\end{itemize}
\end{frame}

\begin{frame}{Partnerships, Data Sharing, and Ecosystems}
	\vspace{15pt}
	\begin{itemize}
		\item Value creation often grows through \textbf{collaboration}.  
		\item \textbf{Strategic partnerships} $\rightarrow$ telcos + city govs use mobility data for traffic planning.  
		\item \textbf{Data sharing} $\rightarrow$ hospital consortia pool patient data for AI-driven diagnosis.  
		\item \textbf{Data ecosystems} $\rightarrow$ e-commerce platforms integrate sellers, buyers, logistics, and payments.  
		\item \textbf{Challenges} $\rightarrow$ privacy, security, fair value distribution demand strong governance.  
		\item Ecosystem-based value $\rightarrow$ network effects boost innovation and digital growth.  
	\end{itemize}
\end{frame}

\section{Monetization at Each Maturity Level (TDWI)}

\begin{frame}{\hfill}
	\centering
	\Huge{\textbf{Monetization at Each Maturity Level (TDWI)}}
\end{frame}

\begin{frame}{Nascent $\rightarrow$ Early $\rightarrow$ Established}
	\vspace{12pt}
	\begin{itemize}
		\item \textbf{Nascent} $\rightarrow$ Data remains fragmented in silos, used mainly for basic reporting. Value limited to simple internal efficiency, no real monetization.  
		\item \textbf{Early} $\rightarrow$ Basic analytics infrastructure is built. Monetization limited to cost savings, inventory optimization, simple analytics. Some organizations experiment with sharing aggregated datasets.  
		\item \textbf{Established} $\rightarrow$ Data becomes integrated across units. Predictive analytics applied in core processes. Monetization expands to dataset licensing, subscription dashboards, internal analytics services.  
	\end{itemize}
\end{frame}

\begin{frame}{Mature $\rightarrow$ Advanced}
	\vspace{12pt}
	\begin{itemize}
		\item \textbf{Mature} $\rightarrow$ Data-driven culture established, AI/ML usage widespread. Monetization more diverse: DaaS, data marketplaces, analytics subscriptions, data-driven advertising.  
		\item \textbf{Advanced} $\rightarrow$ Data becomes the core of the business model. Strong governance and controlled digital ecosystems. Monetization fully diversified: global licensing, cross-industry marketplaces, hybrid models (AI-as-a-Service + data bundling).  
		\item At this stage, data is a strategic capability that is hard to replicate, serving as a primary source of competitive advantage.  
	\end{itemize}
\end{frame}


\section{Conclusion}

\begin{frame}{\hfill}
	\centering
	\Huge{\textbf{Conclusion}}
\end{frame}

\begin{frame}{Conclusion}
	\vspace{12pt}
	\begin{itemize}
		\item Data has transformed into a strategic asset driving business growth, efficiency, and innovation.  
		\item Value realized directly (products/services), indirectly (efficiency gains), and strategically (sustainable competitive advantage).  
		\item The \textit{Big Data Value Chain} shows value is built through systematic stages from acquisition to utilization.  
		\item Monetization models (subscription, advertising, DaaS, marketplace, hybrid) offer diverse opportunities, but sustainability depends on governance, trust, and ethics.  
		\item Organizations that treat data as a core capability are better positioned to lead in digital ecosystems.  
	\end{itemize}
\end{frame}


\end{document}
