\documentclass[aspectratio=169, table]{beamer}

\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, fit}

\usetheme{Pradita}

\title{\Huge Introduction to\\
	\vspace{10pt}
	Big Data for Business}
\subtitle{IT140704 - Big Data for Business}
%\date[Serial]{Penggunaan Large Language Model untuk Pengajaran}
\author{\textbf{Alfa Yohannis}}
\begin{document}
	
	\frame{\titlepage}
	
	
	\begin{frame}[fragile]
		\frametitle{Materi 1}
		\vspace{20pt}
		\begin{columns}[t]
			\column{0.5\textwidth}
			\tableofcontents[sections={1-3}]
			
			\column{0.5\textwidth}
			\tableofcontents[sections={4-6}]
		\end{columns}
	\end{frame}



\section{Introduction}

\begin{frame}[fragile]{Background: The Role of Big Data}
	
	\begin{itemize}
		\item In today’s digital transformation era, data is a strategic asset.
		\item Rapidly growing volumes from internal transactions, customer interactions, and social media create both opportunities and challenges for evidence-based decisions.

		\item Executives must focus on strategic use of data to drive:
		\begin{itemize}
			\item Business value
			\item Operational efficiency
			\item Customer insight
			\item Innovation
		\end{itemize}
	\end{itemize}
\end{frame}


\section{Learning Objectives}

\begin{frame}[fragile]{Course Objectives}
	
	This course equips postgraduate students in business and management with strategic and practical big data knowledge — no programming background required.
	
	\vspace{4pt}
	\begin{itemize}
		\item Understand big data concepts and strategic roles.
		\item Explain big data architecture and infrastructure.
		\item Use visual tools (Power BI, Orange, KNIME) for data exploration and prediction.
		\item Analyse real-world applications across industries.
		\item Identify challenges in ethics, governance, and data monetisation.
	\end{itemize}
\end{frame}

\section{Course Scope}

\begin{frame}[fragile]{Course Scope}
	
	The course covers four key learning areas:
	
	\vspace{4pt}
	\begin{enumerate}
		\item \textbf{Big Data Foundations \& Strategy:} Concepts, 5Vs, Big Data Value Chain, Maturity Models, Architecture.
		\item \textbf{Data Management:} Basics of SQL/NoSQL, data cleaning, source integration.
		\item \textbf{Analytics \& Decision-Making:} BI tools, customer segmentation, prediction, sentiment analysis.
		\item \textbf{Value, Ethics \& Governance:} Data value, policy and compliance.
	\end{enumerate}
\end{frame}


\section{Definition and Characteristics of Big Data}

\begin{frame}[fragile]{Big Data as a New Paradigm}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.5\textwidth}
		\textbf{The Digital Revolution:}
		\begin{itemize}
			\item Data comes from business transactions, IoT sensors, social media, and internal systems.
			\item Data volume is massive and growing rapidly.
		\end{itemize}
		
		\textbf{The Emergence of Big Data:}
		\begin{itemize}
			\item Not just a technical term.
			\item A new paradigm for understanding and managing data.
			\item Focused on creating strategic value from data.
		\end{itemize}
		
		\column{0.5\textwidth}
		\textbf{Chapter Objectives:}
		\begin{itemize}
			\item Introduce the fundamental concept of big data.
			\item Explain its unique characteristics.
			\item Analyse its strategic role in business.
			\item Explore challenges and opportunities in the digital era.
		\end{itemize}
		
		\textbf{Business Relevance:}
		\begin{itemize}
			\item Organisations must be adaptive and data-driven.
			\item Big data is a source of competitive advantage.
		\end{itemize}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Definition and 5V Characteristics of Big Data}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.5\textwidth}
		\textbf{Definition:}
		\begin{itemize}
			\item Large, fast, and varied data that exceeds the capacity of conventional systems.
			\item Known through Gartner's 3V model, later extended to 5V.
			\item Supported by NIST, IEEE P2301, and ISO/IEC 20547-1 definitions.
		\end{itemize}
		
		\textbf{Core Characteristics:}
		\begin{itemize}
			\item \textbf{Volume:} Massive data scale.
			\item \textbf{Velocity:} High-speed data inflow.
			\item \textbf{Variety:} Diverse data formats.
		\end{itemize}
		
		\column{0.5\textwidth}
		\textbf{Additional Characteristics:}
		\begin{itemize}
			\item \textbf{Veracity:} Data quality and reliability.
			\item \textbf{Value:} Strategic insights from data.
		\end{itemize}
		
		\textbf{Business Implications:}
		\begin{itemize}
			\item Organisations require effective data management strategies.
			\item Leveraging the 5V helps achieve competitive advantage.
		\end{itemize}
	\end{columns}
\end{frame}


\section{Strategic Role of Big Data in Organisations}

\begin{frame}[fragile]{Strategic Role of Big Data in Organisations}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.55\textwidth}
		\textbf{Overview:}
		\begin{itemize}
			\item Accelerates digital transformation and innovation.
			\item Delivers insights in real-time and over time.
			\item Data-driven firms perform better.
		\end{itemize}
		
		\textbf{Key Roles:}
		\begin{enumerate}
			\item \textbf{Smarter Decisions:} Actions based on data, not assumptions.
			\item \textbf{Operational Efficiency:} Detects waste, cuts costs, automates tasks.
			\item \textbf{Customer Insight:} Enables personalisation and trend prediction.
		\end{enumerate}
		
		\column{0.45\textwidth}
		\begin{enumerate}
			\setcounter{enumi}{3}
			\item \textbf{Innovation:} Drives new products and models like subscriptions.
			\item \textbf{Competitive Edge:} Improves market speed and differentiation.
		\end{enumerate}
		
		\textbf{Organisational Need:} Requires governance, infrastructure, and culture.  
		\textbf{Conclusion:} Big data is a strategic enabler, beyond just technology.
	\end{columns}
\end{frame}


\section{Evolution of Technology and Digital Business Context}

\begin{frame}[fragile]{Technology Evolution and Digital Business}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.5\textwidth}
		\textbf{Context:}  
		ICT growth has transformed business. Organisations now rely on global connectivity, automation, and digital platforms. Big data supports fast, informed decisions and innovation.
		
		\vspace{6pt}
		\textbf{Tech Milestones:}
		\begin{itemize}
			\item \textbf{Warehouse Era (1980–2000):} Focused on structured historical data using data warehouses.
			\item \textbf{Web \& Cloud Era (2000–2010):} Web apps, cloud, and sensors increased data volume and speed.
		\end{itemize}
		
		\column{0.5\textwidth}
		\begin{itemize}
			\item \textbf{Big Data Era (2010–now):} Tools like Hadoop and Spark enabled distributed large-scale processing.
			\item \textbf{AI \& Real-Time Era (2018–now):} Integration of AI and real-time analytics powers intelligent systems.
		\end{itemize}
		
		\vspace{4pt}
		\textbf{Modern Strategy:}  
		Data is a core business asset—used in supply chains, digital marketing, personalisation, and planning.
		
		\textbf{Conclusion:}  
		Success in the data economy requires infrastructure, analytics skills, governance, and a data-driven mindset.
	\end{columns}
\end{frame}

\section{Challenges and Opportunities in Leveraging Big Data}

\begin{frame}[fragile]{Challenges in Leveraging Big Data}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.5\textwidth}
		\textbf{Key Challenges:}
		\begin{enumerate}
			\item \textbf{Data Quality \& Integration:} Inconsistent formats and incomplete data from various sources complicate analysis.
			\item \textbf{Talent Gap:} Limited availability of professionals with advanced data analytics skills.
			\item \textbf{Unclear Data Strategy:} Lack of alignment between data initiatives and business objectives.
		\end{enumerate}
		
		\column{0.5\textwidth}
		\begin{enumerate}
			\setcounter{enumi}{3}
			\item \textbf{Security \& Privacy Risks:} High volumes of sensitive data increase exposure to leaks and misuse.
			\item \textbf{Governance \& Ethics Issues:} Unclear roles and responsibilities for data accuracy, usage, and retention.
		\end{enumerate}
		
		\textbf{Conclusion:}  
		Addressing these challenges requires integrated efforts across people, processes, and technology to support secure and meaningful data use.
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Strategic Opportunities from Big Data}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.5\textwidth}
		\textbf{Opportunities:}
		\begin{itemize}
			\item \textbf{Operational Efficiency:} Data-driven forecasting and automation improve resource allocation and supply chains.
			\item \textbf{Deeper Customer Insight:} Enables personalised marketing and customer engagement based on behaviour and feedback.
			\item \textbf{Product \& Service Innovation:} Market trends and data insights drive new solution development.
		\end{itemize}
		
		\column{0.5\textwidth}
		\begin{itemize}
			\item \textbf{Real-Time Decision Making:} Integrated dashboards and AI enable instant strategic and operational responses.
			\item \textbf{Data Monetisation:} Internal data becomes a valuable asset through subscription models and inter-organisational sharing.
		\end{itemize}
		
		\textbf{Conclusion:}  
		Organisations must align technology, talent, processes, and governance to reduce risks and fully realise big data’s strategic value.
	\end{columns}
\end{frame}

\section{Examples of Big Data Applications in Business}

\begin{frame}[fragile]{Big Data Use Cases Across Industries}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.5\textwidth}
		\begin{itemize}
			\item \textbf{Retail \& E-commerce:}  
			Customer behaviour analysis, product recommendations, and supply chain optimisation.
			
			\item \textbf{Finance \& Banking:}  
			Fraud detection, credit risk scoring, and personalised financial services.
			
			\item \textbf{Manufacturing \& Industry 4.0:}  
			IoT and predictive analytics for maintenance and smart factory operations.
		\end{itemize}
		
		\column{0.5\textwidth}
		\begin{itemize}
			\item \textbf{Transport \& Logistics:}  
			Route and delivery optimisation based on real-time demand and location data.
			
			\item \textbf{Healthcare:}  
			Medical record analytics for early diagnosis and personalised treatment.
			
			\item \textbf{Education:}  
			Learning analytics to monitor student progress and adapt instruction.
		\end{itemize}
		
		\textbf{Conclusion:}  
		Big data enables innovation and competitive advantage when supported by the right infrastructure and capabilities.
	\end{columns}
\end{frame}

\section{Conclusion}

\begin{frame}[fragile]{Conclusion}
	\vspace{20pt}
	\small
	\begin{columns}[T]
		\column{0.62\textwidth}
		\textbf{Key Takeaways:}
		\begin{itemize}
			\item Big data is defined by five core dimensions: volume, velocity, variety, veracity, and value.
			\item It requires new approaches in data collection, storage, and analysis beyond conventional systems.
			\item Beyond technology, big data reshapes business models, decision-making, and organisational strategies.
		\end{itemize}
		
		\textbf{Strategic Value:}
		\begin{itemize}
			\item Enables competitive advantage through efficiency and innovation.
			\item Supported by technologies like cloud computing, IoT, and AI.
		\end{itemize}
		
		\column{0.38\textwidth}
		\textbf{Challenges Ahead:}
		\begin{itemize}
			\item Governance, security, ethics, and human resource readiness remain critical concerns.
		\end{itemize}
		
		\textbf{Sector Applications:}
		\begin{itemize}
			\item Value creation depends on alignment between technology, business strategy, and data culture.
		\end{itemize}
		
		\textbf{Next Sessions:}  
		Will explore data lifecycle management, infrastructure selection, and practical analytics across industries.
	\end{columns}
\end{frame}

\section{Course Plan: Big Data for Business}

\begin{frame}[fragile]{Course Topics and Focus (Weeks 1–5)}
	\vspace{20pt}
	\small
	\begin{enumerate}
		\item \textbf{Introduction to Big Data in Business} — Key definitions, business relevance, and case illustrations.
		\item \textbf{Big Data Strategy \& Frameworks} — Overview of 5Vs, Big Data Value Chain, BDMM, and DAMA.
		\item \textbf{Big Data Infrastructure \& Technologies} — Technologies such as Hadoop, Spark, Kafka, NoSQL, and cloud platforms.
		\item \textbf{Understanding Databases \& SQL} — Structured data basics and SQL querying using online tools.
		\item \textbf{NoSQL \& Semi-Structured Data} — JSON and MongoDB for flexible schema and document-based data.
	\end{enumerate}
\end{frame}

\begin{frame}[fragile]{Course Topics and Focus (Weeks 6–10)}
	\vspace{20pt}
	\small
	\begin{enumerate}
		\setcounter{enumi}{5}
		\item \textbf{Data Cleaning \& Preparation} — Techniques for ensuring quality and consistency; KNIME workflows.
		\item \textbf{Business Intelligence \& Dashboards} — Building dashboards, KPIs, and visual insights with Power BI.
		\item \textbf{Introduction to Machine Learning} — Classification, clustering, and BI comparison using Orange.
		\item \textbf{Customer Segmentation (Clustering)} — Market segmentation using unsupervised learning (k-means).
		\item \textbf{Prediction \& Forecasting} — Forecasting churn, revenue, or risk with regression models.
	\end{enumerate}
\end{frame}

\begin{frame}[fragile]{Course Topics and Focus (Weeks 11–14)}
	\vspace{20pt}
	\small
	\begin{enumerate}
		\setcounter{enumi}{10}
		\item \textbf{Text \& Sentiment Analysis} — NLP for analysing feedback and sentiment using Orange.
		\item \textbf{Data-Driven Decision Frameworks} — CRISP-DM, OSEMN, and converting insights into action.
		\item \textbf{Data Monetisation \& Value Realisation} — Strategies to turn data into economic and strategic value.
		\item \textbf{Ethics, Governance \& Maturity Review} — Data ethics, privacy, governance, and BDMM self-assessment.
	\end{enumerate}
\end{frame}

\begin{frame}[fragile]{Final Assignment Guidelines}
	\vspace{20pt}
	\small
	Students are required to write an academic paper focusing on one of the following themes. The paper should be based on a structured approach and, where possible, include practical application or evaluation.
	
	\begin{enumerate}
		\item \textbf{Systematic Literature Review on the Data Value Chain}.  
		Review existing models of the Data Value Chain (DVC), summarise key steps, and propose a refined version. Apply this framework to a small-scale case and evaluate the usefulness of the framework.
		
		\item \textbf{Insight Discovery from Captured Data}.  
		Collect and analyse secondary data or internal workplace data to discover new business insights. Highlight the data preparation, analysis process, and conclusions.
		
		\item \textbf{Maturity Assessment and Improvement Strategy}.  
		Evaluate an organisation's big data maturity using a recognised framework (e.g., BDMM). Recommend actionable tactics to improve maturity, implement them, and evaluate the results.
		
		\item \textbf{Other Topics Related to Big Data}.  
		Proposals on other relevant topics in big data are welcome.
	\end{enumerate}
\end{frame}



	
%\section{Bab 1: Pengenalan ChatGPT}
%\begin{frame}{\hfill}
%	\centering
%	\Huge{\textbf{Bab 1: Pengenalan ChatGPT}}
%\end{frame}	
		\end{document}
