\documentclass[aspectratio=169, table]{beamer}

\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, fit}

\usetheme{Pradita}

\title{\Huge Strategy and Big Data\\
	\vspace{10pt}
	Frameworks}
\subtitle{IT140704 - Big Data for Business}
%\date[Serial]{Penggunaan Large Language Model untuk Pengajaran}
\author{\textbf{Alfa Yohannis}}
\begin{document}
	
	\frame{\titlepage}
	
	
	\begin{frame}[fragile]
		\frametitle{Materi 1}
		\vspace{20pt}
		\begin{columns}[t]
			\column{0.5\textwidth}
			\tableofcontents[sections={1-3}]
			
			\column{0.5\textwidth}
			\tableofcontents[sections={4-6}]
		\end{columns}
	\end{frame}

	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Why do you think so many big data projects fail despite having advanced technology? What else might be missing?}}
	\end{frame}
	
	\section{Big Data Strategy and Frameworks}
	
	\begin{frame}[fragile]{Why Strategy Matters in Big Data}
		\vspace{20pt}
		\small
		\begin{columns}[T]
			\column{0.5\textwidth}
			\textbf{Strategic Importance:}
			\begin{itemize}
				\item Aligns data with business goals.
				\item Avoids fragmented and costly initiatives.
				\item Supports governance and performance.
			\end{itemize}
			
			\textbf{Common Failures:} Over 85\% of big data projects fail due to lack of alignment between technology, process, and capability.
			
			
			\vspace{5pt}
			\textbf{Big Data Requires Strategy:}
			\begin{itemize}
				\item Business-driven
				\item Measurable impact
				\item Scalable and secure
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Covered Frameworks:}
			\begin{itemize}
				\item \textbf{5Vs Model:} Core characteristics of big data.
				\item \textbf{Big Data Value Chain :} Step-by-step data lifecycle.
				\item \textbf{Big Data Maturity Model (BDMM):} Assessing organisational readiness.
				\item \textbf{DAMA-DMBOK:} Governance, quality, stewardship.
			\end{itemize}
			
			\textbf{Strategic Outcome:}  
			Big data strategy bridges vision and execution, supporting digital transformation and data-driven decision-making.
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Key Elements of a Big Data Strategy}
		\vspace{20pt}
		\small
		\begin{columns}[T]
			\column{0.5\textwidth}
			\begin{enumerate}
				\item \textbf{Business Vision:}  
				Start from business priorities like customer experience or innovation.
				
				\item \textbf{Architecture \& Technology:}  
				Select platforms matching data scale, speed, and integration needs.
				
				\item \textbf{Governance \& Compliance:}  
				Ensure ethical, secure, and rule-compliant data use.
			\end{enumerate}
			
			\column{0.5\textwidth}
			\begin{enumerate}
				\setcounter{enumi}{3}
				\item \textbf{Organisational Capability:}  
				Invest in skills, data literacy, and culture.
				
				\item \textbf{Metrics \& Impact:}  
				Set KPIs to measure operational, financial, and strategic outcomes.
			\end{enumerate}
			
			\vspace{6pt}
			\textbf{Conclusion:}  
			An effective strategy integrates technology, people, and governance to ensure data delivers real business value.
		\end{columns}
	\end{frame}
	
	
	\section{5Vs Conceptual Framework for Big Data}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{5Vs Conceptual Framework for Big Data}}
	\end{frame}
	
	\begin{frame}[fragile]{Introduction and Volume}
		\vspace{20pt}
		\textbf{Introduction:} The 5Vs framework is widely used to conceptualise big data characteristics. Originally introduced by Doug Laney (2001) with Volume, Velocity, and Variety, it has expanded to include Veracity and Value.
		
		\vspace{10pt}
		\begin{columns}[T]
			\column{0.5\textwidth}
			\textbf{1. Volume:}
			\begin{itemize}
				\item Refers to the massive amount of data generated from various sources.
				\item Requires scalable distributed storage (e.g., Hadoop HDFS, cloud).
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Examples:}
			\begin{itemize}
				\item Amazon handles billions of daily product interactions and transactions.
				\item Hospitals store high-volume radiology, lab, and patient records.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Velocity and Variety}
		\vspace{20pt}
		\begin{columns}[T]
			\column{0.5\textwidth}
			\textbf{2. Velocity:}
			\begin{itemize}
				\item Refers to the speed of data generation and processing.
				\item Enables real-time analytics using Apache Kafka, Spark Streaming.
			\end{itemize}
			
			\vspace{10pt}
			\textbf{Examples:}
			\begin{itemize}
				\item Gojek/Grab use live GPS data for fleet dispatch.
				\item Stock trading systems respond to data within milliseconds.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{3. Variety:}
			\begin{itemize}
				\item Represents the diversity of data types (structured, semi, unstructured).
				\item Integration across formats enhances business insights.
			\end{itemize}
			
			\vspace{10pt}
			\textbf{Examples:}
			\begin{itemize}
				\item TikTok processes videos, audio, and user metadata.
				\item Telkom combines billing and call centre transcripts.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Veracity and Value}
		\vspace{20pt}
		\begin{columns}[T]
			\column{0.5\textwidth}
			\textbf{4. Veracity:}
			\begin{itemize}
				\item Describes data quality, consistency, and reliability.
				\item Requires cleansing, validation, and deduplication.
			\end{itemize}
			
			\vspace{10pt}
			\textbf{Examples:}
			\begin{itemize}
				\item Medical input errors risk misdiagnosis.
				\item Banking records may cause credit scoring mistakes if duplicated.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{5. Value:}
			\begin{itemize}
				\item Focuses on extracting business value from analytics.
				\item Value includes efficiency, savings, innovation.
			\end{itemize}
			
			\vspace{10pt}
			\textbf{Examples:}
			\begin{itemize}
				\item Netflix boosts retention with personal recommendations.
				\item DHL uses predictive models to cut costs and emissions.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	
	\section{Big Data Value Chain}
	
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Big Data Value Chain}}
	\end{frame}
	
	
	\begin{frame}[fragile]{Overview: Big Data Value Chain}
		\vspace{20pt}
		\begin{columns}[T]
			\column{0.6\textwidth}
			\begin{itemize}
				\item A strategic framework that maps the full lifecycle of data use — from initial planning to capturing, transforming, analysing, and generating business impact.
				\item Ensures that data activities are aligned with business objectives and measured for effectiveness.
				\item Consists of 9 stages: \textbf{Define, Capture, Curate, Process, Analyse, Visualise, Decide, Act, Measure}.
				\item Emphasises both the technical and business dimensions of big data initiatives.
			\end{itemize}
			
			\column{0.4\textwidth}
			\begin{enumerate}
				\item \textbf{Define} — Set goals, data needs, and analysis questions.\\
				\textit{e.g.:} Define focus, needed data, tools, methods.
			\end{enumerate}
		\end{columns}
	\end{frame}
	
	
	\begin{frame}[fragile]{Big Data Value Chain: Capture to Analyse}
		\vspace{20pt}
		\begin{columns}[T]
			\column{0.5\textwidth}
			\begin{enumerate}
				\setcounter{enumi}{1}
				
				\item \textbf{Capture} — Collect structured/unstructured data from various sources.\\
				\textit{e.g.:} IoT sensors, transactions, logs.\\
				Tools: API, web scraping, Apache NiFi.
				
				\item \textbf{Curate} — Clean, validate, standardise, integrate data.\\
				\textit{e.g.:} De-duplication, harmonisation.\\
				Tools: KNIME, Talend, Alteryx.
			\end{enumerate}
			
			\column{0.5\textwidth}
			\begin{enumerate}
				\setcounter{enumi}{3}
				
				
				
				\item \textbf{Process} — Transform and store data via ETL, batch, or real-time processing.\\
				\textit{e.g.:} Daily Hadoop jobs, Spark streaming.\\
				Tools: Hadoop, Spark, Airflow.
				
				\item \textbf{Analyse} — Explore and model data using statistical or ML techniques.\\
				\textit{e.g.:} Customer segmentation, churn prediction.\\
				Tools: Orange, scikit-learn, R, Python.
			\end{enumerate}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Big Data Value Chain: Visualise to Measure}
		\vspace{20pt}
		\begin{columns}[t]
			\column{0.5\textwidth}
			\begin{enumerate}
				\setcounter{enumi}{5}
				
				
				\item \textbf{Visualise} — Present findings with charts, dashboards, and visuals.\\
				\textit{e.g.:} BI dashboards, Tableau clusters.
				
				\item \textbf{Decide} — Make informed decisions (operational to strategic).\\
				\textit{e.g.:} Dynamic pricing, campaign priorities.
			\end{enumerate}
			
			\column{0.5\textwidth}
			\begin{enumerate}
				\setcounter{enumi}{7}
				\item \textbf{Act} — Implement actions based on insights into real-world business processes.\\
				\textit{e.g.:} Trigger campaign, adjust operations.
				
				\item \textbf{Measure} — Evaluate outcomes to assess value, improve strategy.\\
				\textit{e.g.:} ROI analysis, A/B testing, KPI tracking.
			\end{enumerate}
		\end{columns}
	\end{frame}
	
	
	\begin{frame}[fragile]{Applications of Big Data Value Chain (1)}
		\vspace{20pt}
		\small
		\begin{columns}[t]
			\column{0.5\textwidth}
			\textbf{Retail and E-commerce}
			\begin{itemize}
				\item \textbf{Define:} Improve conversions via personalisation.
				\item \textbf{Capture:} Collect click and purchase logs.
				\item \textbf{Curate:} Clean bots, fix session errors.
				\item \textbf{Process:} Load into warehouse, run ETL.
				\item \textbf{Analyse:} Segment users, detect patterns.
				\item \textbf{Visualise:} Product recommendation dashboards.
				\item \textbf{Decide:} Personalise homepage content.
				\item \textbf{Act:} Apply dynamic deals and feeds.
				\item \textbf{Measure:} Track CTR, cart size, sales.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Manufacturing}
			\begin{itemize}
				\item \textbf{Define:} Reduce downtime with predictions.
				\item \textbf{Capture:} Stream IoT sensor data.
				\item \textbf{Curate:} Filter outliers, standardise values.
				\item \textbf{Process:} Store in time-series DB.
				\item \textbf{Analyse:} Predict failure via ML models.
				\item \textbf{Visualise:} Show equipment health status.
				\item \textbf{Decide:} Notify preventive maintenance.
				\item \textbf{Act:} Auto-schedule technician visits.
				\item \textbf{Measure:} Downtime and cost reduction.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Applications of Big Data Value Chain (2)}
		\vspace{20pt}
		\small
		\begin{columns}[t]
			\column{0.5\textwidth}
			\textbf{Finance and Banking}
			\begin{itemize}
				\item \textbf{Define:} Detect transaction fraud quickly.
				\item \textbf{Capture:} Ingest streaming payments.
				\item \textbf{Curate:} Fix errors, remove duplicates.
				\item \textbf{Process:} Stream via Kafka/Flink.
				\item \textbf{Analyse:} Spot fraud with ML models.
				\item \textbf{Visualise:} Alert dashboards for analysts.
				\item \textbf{Decide:} Auto-flag risky actions.
				\item \textbf{Act:} Block or freeze accounts.
				\item \textbf{Measure:} Fraud drop, accuracy rate.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Public Sector}
			\begin{itemize}
				\item \textbf{Define:} Evaluate policy and services.
				\item \textbf{Capture:} Pull survey and service data.
				\item \textbf{Curate:} Merge census and usage logs.
				\item \textbf{Process:} Aggregate by region or type.
				\item \textbf{Analyse:} Find service gaps, trends.
				\item \textbf{Visualise:} Dashboards for decision makers.
				\item \textbf{Decide:} Recommend targeted policies.
				\item \textbf{Act:} Roll out public programmes.
				\item \textbf{Measure:} Assess impact via KPIs.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\section{Big Data Maturity Model (BDMM)}
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Big Data Maturity Model (BDMM)}}
	\end{frame}

	\begin{frame}[fragile]{BDMM: Definition and Purpose}
		
		\begin{itemize}
			\item The Big Data Maturity Model (BDMM) is a strategic framework to assess an organisation's readiness and capability in managing and leveraging big data.
			\item It maps current conditions (\textit{as-is}) and future targets (\textit{to-be}) through progressive stages.
			\item BDMM provides a roadmap for becoming data-driven and aligns with business transformation initiatives.
		\end{itemize}
	\end{frame}
	
	\begin{frame}[fragile]{BDMM Levels (1--2)}
		\textbf{Level 1: Initial}
		\begin{itemize}
			\item Also known as \textit{Nascent} or \textit{Ad-hoc}.
			\item No formal data strategy; isolated efforts.
			\item \textit{Example:} Marketing uses Excel for surveys; finance stores data locally.
		\end{itemize}
		
		\textbf{Level 2: Managed}
		\begin{itemize}
			\item Early experiments or pilots in isolated units.
			\item Governance is minimal.
			\item \textit{Example:} Logistics uses Power BI dashboard; HR runs churn models independently.
		\end{itemize}
	\end{frame}
	
	\begin{frame}[fragile]{BDMM Levels (3--5)}
		
		\textbf{Level 3: Defined}
		\begin{itemize}
			\item Documented and repeatable practices; shared infrastructure.
			\item \textit{Example:} Central data warehouse; demand forecasting for inventory.
		\end{itemize}
		
		\textbf{Level 4: Integrated}
		\begin{itemize}
			\item Cross-functional adoption with strong governance.
			\item \textit{Example:} Shared KPI dashboards; governance committee in place.
		\end{itemize}
		
		\textbf{Level 5: Optimized}
		\begin{itemize}
			\item Predictive and real-time decision-making embedded.
			\item \textit{Example:} Auto product recommendations; simulation-based business strategy.
		\end{itemize}
	\end{frame}
	
	\begin{frame}[fragile]{Assessment Dimensions}
		\vspace{18pt}
		\begin{itemize}
			\item \textbf{Strategy and Vision:} Alignment with business objectives. \textit{Example:} Customer insights for service innovation.
			\item \textbf{Data and Governance:} Policies and roles. \textit{Example:} Sensitive data classification; data stewards.
			\item \textbf{Technology:} Infrastructure readiness. \textit{Example:} Cloud-based data lake; Kafka streaming.
			\item \textbf{Analytics:} Advanced tools and use cases. \textit{Example:} Power BI dashboards; churn prediction models.
			\item \textbf{Organisation and HR:} Training and structure. \textit{Example:} Regular analyst training; dedicated data teams.
			\item \textbf{Security and Privacy:} Compliance and risk. \textit{Example:} Role-based access control; PDP/GDPR adherence.
			\item Indicators support self-assessment via surveys, benchmarks, and workshops.
		\end{itemize}
	\end{frame}

	
	\begin{frame}[fragile]{Using BDMM Strategically}
		\vspace{20pt}
		\begin{itemize}
			\item \textbf{Gap Analysis:} Identify current level and priority areas. \textit{Example:} Customer service unit operates in silo.
			\item \textbf{Roadmap Planning:} Transition roadmap to higher maturity. \textit{Example:} Move from opportunistic to systematic with a data warehouse.
			\item \textbf{Investment Planning:} Align investments to capability gaps. \textit{Example:} Adopt integration platforms to improve quality and speed.
			\item \textbf{Cross-Functional Alignment:} Foster collaboration across departments.
			\item \textbf{Impact Measurement:} Monitor strategic outcomes like ROI and time savings.
		\end{itemize}
	\end{frame}
	
		\section{Data Management Body of Knowledge}
	\begin{frame}{\hfill}
		\centering
		\Huge{\textbf{Data Management Body of Knowledge}}
		
		\vspace{20pt}
		\large{\textbf{Data Management Association International}}
	\end{frame}
	
	\begin{frame}[fragile]{DAMA-DMBOK: Overview and Domains}
		\vspace{20pt}
		\begin{columns}
			\column{0.5\textwidth}
			\textbf{What is DAMA-DMBOK?}
			\begin{itemize}
				\item A comprehensive framework for data governance developed by DAMA International.
				\item Covers principles, roles, and best practices across the data lifecycle.
				\item Technology-neutral, focusing on policies and organisational capabilities.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Key Domains (1–6):}
			\begin{itemize}
				\item \textbf{Data Governance}
				\item \textbf{Data Architecture}
				\item \textbf{Data Modelling \& Design}
				\item \textbf{Data Storage \& Operations}
				\item \textbf{Data Security}
				\item \textbf{Data Integration \& Interoperability}
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{DAMA-DMBOK: Domains (7–11)}
		\vspace{20pt}
		\begin{columns}
			\column{0.5\textwidth}
			\textbf{Key Domains (7–11):}
			\begin{itemize}
				\item \textbf{Document \& Content Management}
				\item \textbf{Reference \& Master Data}
				\item \textbf{Data Warehousing \& BI}
				\item \textbf{Metadata Management}
				\item \textbf{Data Quality Management}
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Integration:}
			\begin{itemize}
				\item All domains are interconnected.
				\item Foundation for sustainable, scalable data governance.
				\item Aligns well with Big Data strategy, maturity models, and compliance frameworks.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Strategic Contributions of DAMA}
		\vspace{20pt}
		\begin{columns}
			\column{0.5\textwidth}
			\textbf{Supports Long-term Strategy}
			\begin{itemize}
				\item Provides clear roles (e.g., data steward).
				\item Aligns governance with transformation roadmaps.
			\end{itemize}
			
			\textbf{Ensures Compliance and Ethics}
			\begin{itemize}
				\item GDPR, UU PDP, and internal audit alignment.
				\item Consent and access control based on defined policies.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Enables Quality and Consistency}
			\begin{itemize}
				\item DQ and metadata as performance indicators.
				\item Uniform data definitions organisation-wide.
			\end{itemize}
			
			\textbf{Bridges Tech and Policy}
			\begin{itemize}
				\item Co-owned validation rules by engineers and business users.
				\item Collaborative cross-division governance.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\begin{frame}[fragile]{Practical Organisational Implications}
		\vspace{20pt}
		\begin{columns}
			\column{0.5\textwidth}
			\textbf{Role Clarity and Accountability}
			\begin{itemize}
				\item Define data owner, steward, custodian roles.
				\item Required coordination for migrations and audits.
			\end{itemize}
			
			\textbf{Transparency and Auditability}
			\begin{itemize}
				\item Lineage metadata supports traceability.
				\item KPI definitions avoid misinterpretation.
			\end{itemize}
			
			\column{0.5\textwidth}
			\textbf{Risk and Cost Reduction}
			\begin{itemize}
				\item Eliminate redundancy and unauthorised access.
				\item Role-based authentication and 30\% storage savings.
			\end{itemize}
			
			\textbf{Analytics Scalability}
			\begin{itemize}
				\item AI/ML teams use official data catalogue.
				\item Cloud platforms can be easily integrated because of standardisation and documentation.
			\end{itemize}
		\end{columns}
	\end{frame}
	
	\section{Conclusion}
	
	\begin{frame}[fragile]{Conclusion}
		\small
		Big data strategies and frameworks are essential for turning data into lasting business value. This chapter introduced several core conceptual and practical approaches:
		
		\vspace{6pt}
		\begin{itemize}
			\item \textbf{5Vs:} Capturing data's core traits — volume, velocity, variety, veracity, and value.
			\item \textbf{Big Data Value Chain:} Outlining end-to-end data flow from goals to decisions.
			\item \textbf{BDMM:} Enabling maturity assessment and roadmap for data capability growth.
			\item \textbf{DAMA-DMBOK:} Structuring governance roles and policies for reliable data use.
		\end{itemize}
		
		\vspace{6pt}
		Together, these frameworks bridge technology, organisation, people, and policy—crucial for a successful, data-driven transformation.
	\end{frame}
	
	
	
	
	%\section{Bab 1: Pengenalan ChatGPT}
	%\begin{frame}{\hfill}
	%	\centering
	%	\Huge{\textbf{Bab 1: Pengenalan ChatGPT}}
	%\end{frame}	
\end{document}
